---
title: "Adjusted and Unadjusted Analyses"
output: html_document
---

```{r, echo = F, warning=F, message = F}
library(sjPlot)
library(sjstats)
library(tableone)
library(rcompanion)

model_info <- function(fit){
  model_sum <- summary(fit)
  odds_ratio <- cbind(OR = exp(fit$coef), exp(confint(fit)))
  my_list <- list(model_sum, odds_ratio)
  names(my_list) <- c("Model Summary", "OR Summary")
  return(my_list)
}

## to_hours
## convert MIMIC-III date time format to hours
to_hours <- function(some_time){
  return(as.numeric(strptime(some_time, "%Y-%m-%d %H:%M:%S", tz = "EST"))/(60*60))
}

## Factor data which have a few levels
factor_data <- function(dat){
    ## Any row with levels less than the log of the number of rows
    cols_to_factor <- sapply(dat, function(col) length(unique(col)) < log10(length(col) * 1000) )
    dat[cols_to_factor] <- lapply(dat[cols_to_factor], factor)
    return(dat)
}


plotDat <- function(dat, column, x_col, bs, mn, xl, yl){
  tmp <- as.matrix(table(dat[[column]], dat[[x_col]]))
  prop <- prop.table(tmp, margin = 2)#2 for column-wise proportions
  par(mar = c(5.0, 4.0, 4.0, 15), xpd = TRUE)
  barplot(prop, col = c("#808080", "#000000"), beside = bs, width = 2, main = mn, xlab = xl, ylab = yl)
  legend("topright", inset = c(-0.25,0), fill = c("#808080", "#000000"), legend=rownames(prop))
}


#plotDat(temp, "NQF", "ADMISSION_TYPE", F, "Admission Type", "Admission Type", "Frequency")
#test <- table(temp$ADMISSION_TYPE, temp$NQF)
#test
#chisq.test(test)
##pairwiseNominalIndependence(
#  as.matrix(test), 
#  fisher = T, gtest = F, chisq = F, method = "fdr")
```

### Load data

```{r}
dat <- read.csv("~/caregivers/data/processed/regression_data11Jun2020.csv", 
                header = T, stringsAsFactors = F)

## Drop columns
dat$DURATION_HOURS <- NULL
dat$DURATION_DAYS <- NULL

```

## Unadjusted Results

```{r}
tmp <- dat[,c("SPOUSE_FIRST_48",
                "CHILD_FIRST_48",
                "MORT_1YR",
                "MORT_3MO",
                "AGE",
                "GENDER",
                "LANGUAGE",
                "ETHNICITY",
                "MARITAL_STATUS",
                "SOFA_SCORE",
                "ELIXHAUSER_SCORE",
                "PERC_ICU_VENT",
                "LOS_HOSPITAL",
                "LOS_ICU")]

## Factor
tmp <- factor_data(tmp)

## Define non normal variables
non_normal <- c("AGE", 
                "PERC_ICU_VENT",
                "SOFA_SCORE",
                "ELIXHAUSER_SCORE",
                "LOS_HOSPITAL",
                "LOS_ICU")



overall_tab <- CreateTableOne(data = tmp)

#write.csv(print(overall_tab, non_normal = non_normal, quote = F, noSpaces = T, #showAllLevels = TRUE), file = "~/Desktop/CG_Data/caregivers/overall_table_one12Jun2020.csv")

child_tab <- CreateTableOne(data = tmp, strata = "CHILD_FIRST_48")

#write.csv(print(child_tab, nonnormal = non_normal, 
#                quote = F, noSpaces = T, showAllLevels = T), file = "~/Desktop/CG_Data/caregivers/child_table_one12Jun2020.csv")

spouse_tab <- CreateTableOne(data = tmp, strata = "SPOUSE_FIRST_48")

#write.csv(print(spouse_tab, nonnormal = non_normal, 
#                quote = F, noSpaces = T, showAllLevels = T), file = "~/Desktop/CG_Data/caregivers/Spouse_table_one12Jun2020.csv")
```


## Family Involvement and Ventilator Use

### Subset data

__Outcome:__ `PERC_ICU_VENT`

1. Demographics: `AGE`, `GENDER`, `LANGUAGE`, `MARITAL_STATUS`
2. Clinical: `SOFA`
3. Communication: `SPOUSE_FIRST_48`, `CHILD_FIRST_48`

```{r}
## Subset
temp <- dat[,c("PERC_ICU_VENT",
                "AGE",
                "GENDER",
                "LANGUAGE",
                "MARITAL_STATUS",
                "SOFA_SCORE",
                "ELIXHAUSER_SCORE",
                "SPOUSE_FIRST_48",
                "CHILD_FIRST_48")]


## Bucket Ages
dat$AGE_BUCKET <- ifelse(dat$AGE <=  )

## Factor
temp <- factor_data(temp)
```


### Regression

```{r}
model <- lm(PERC_ICU_VENT ~., data = temp)

summary(model)

tab_model(model)
```

## Family Involvement and Length of Stay

#### Note:

```{r}
hist(dat$LOS_HOSPITAL)

## transform using natural log
dat$LOG_LOS <- log(dat$LOS_HOSPITAL)

hist(dat$LOG_LOS)
```


```{r}
## Subset
temp <- dat[,c("LOS_HOSPITAL",
                "AGE",
                "GENDER",
                "LANGUAGE",
                "MARITAL_STATUS",
                "SOFA_SCORE",
                "SPOUSE_FIRST_48",
                "CHILD_FIRST_48")]

## Factor
temp <- factor_data(temp)
```

### Regression

```{r}
model <- lm(LOS_HOSPITAL ~., data = temp)


tab_model(model)


## Recreate model with log transformation

## Subset
temp <- dat[,c("LOG_LOS",
                "AGE",
                "GENDER",
                "LANGUAGE",
                "MARITAL_STATUS",
                "SOFA_SCORE",
                "SPOUSE_FIRST_48",
                "CHILD_FIRST_48")]

## Factor
temp <- factor_data(temp)


## Recreate model with log transformation
model <- lm(LOG_LOS ~., data = temp)

tab_model(model)


summary(model)



round(cbind((exp(model$coefficients)-1)*100), 2)

```

Only the dependent/response variable is log-transformed. 



Exponentiate the coefficient, subtract one from this number, and multiply by 100. This gives the percent increase (or decrease) in the response for every one-unit increase in the independent variable. Example: the coefficient is 0.198. (exp(0.198) – 1) * 100 = 21.9. For every one-unit increase in the independent variable, our dependent variable increases by about 22%.


https://stats.idre.ucla.edu/sas/faq/how-can-i-interpret-log-transformed-variables-in-terms-of-percent-change-in-linear-regression/

 In such models where the dependent variable has been log-transformed and the predictors have not. To interpet the amount of change in the original metric of the outcome, we first exponentiate the coefficient of census to obtain exp(0.00055773)=1.000558. To calculate the percent change, we can subtract one from this number and multiply by 100. Thus, for a one unit increase in the average daily number of patients (census), the average length of stay (length) increases by 0.06 percent.

## Patient-level factors associated with family involvement

### Child

```{r}

## Subset
temp <- dat[,c("AGE",
                "GENDER",
                "ETHNICITY",
                "LANGUAGE",
                "MARITAL_STATUS",
                "SOFA_SCORE",
                "SPOUSE_FIRST_48",
                "CHILD_FIRST_48")]

## Center scale
center_scale <- function(x) {
    scale(x, scale = FALSE)
}

temp$AGE <- center_scale(temp$AGE)  
  

## Factor
temp <- factor_data(temp)


model <- glm(CHILD_FIRST_48 ~ 
              GENDER +
              AGE +
              ETHNICITY +
              LANGUAGE +
              MARITAL_STATUS +
              SOFA_SCORE +
               SPOUSE_FIRST_48, 
                   data = temp,
                   family = binomial(link = "logit"))

tab_model(model)

```




```{r}
## Subset
temp <- dat[,c("AGE",
                "GENDER",
                "ETHNICITY",
                "LANGUAGE",
                "MARITAL_STATUS",
                "SOFA_SCORE",
                "SPOUSE_FIRST_48",
                "CHILD_FIRST_48")]

temp$AGE <- center_scale(temp$AGE)  
  

## Factor
temp <- factor_data(temp)


model <- glm(SPOUSE_FIRST_48 ~ 
              GENDER +
              AGE +
              ETHNICITY +
              LANGUAGE +
              MARITAL_STATUS +
              SOFA_SCORE +
               CHILD_FIRST_48, 
                   data = temp,
                   family = binomial(link = "logit"))

tab_model(model)
```
