---
title: "Caregivers Dictionary Validation"
output: html_document
---

## Strategy

1. We have looked at an annotation cohort of all ventilated patients' Physician/Nursing notes to create a keyword dictionary optimized on sensitivity

2. These dictionaries, `child` and `spouse`, will now be validated on random, semi-overlapping, 1000-note samples with the purpose of optimizing on specificity

3. These samples focus on the Medical ICU, include Social Work and Case Management notes, but remain focused on the first 48hrs of the hospital admission

## Load Annotation Cohort Set
### Annotation Cohort is Enriched with Concepts

```{r}
## Load Data
dat <- read.csv("~/caregivers/data/first_48_entire_cohort26Jun2020.csv", header = T, stringsAsFactors = F)

## Subset minimum necessary variables: ROW_ID, TEXT
dat <- dat[ ,c("ROW_ID", "TEXT")]

## Randomize notes
## First set seed
set.seed(1337)
dat <- dat[sample(1:nrow(dat)),]

## Write the resulting data
write.csv(dat, file = "~/caregivers/data/keyword_dict_investigation_set26Jun2020.csv", row.names = F)

## Sample 1000 for each person
write.csv(dat[1:1000,], file = "~/Desktop/caregivers annotation/caregivers_notes_1_26Jun2020.csv", row.names = F)

write.csv(dat[900:1899,], file = "~/Desktop/caregivers annotation/caregivers_notes_2_26Jun2020.csv", row.names = F)

write.csv(dat[c(900:1000, 1900:2798),], file = "~/Desktop/caregivers annotation/caregivers_notes_3_26Jun2020.csv", row.names = F)
```

## Dictionaries

Dictionaries focus on key terms which indicate communication with persons in the ICU

```{r, warning=FALSE}
child <- unique(clean_text(readLines("~/caregivers/dictionaries/child12Jun2020.txt"), F))

child <- readLines("~/caregivers/dictionaries/child12Jun2020.txt")

write(child, file = '~/caregivers/dictionaries/child_dict_25Jun2020.txt', sep = ',', ncolumns = length(child))

spouse <- readLines("~/caregivers/dictionaries/spouse12Jun2020.txt")

write(spouse, file = '~/caregivers/dictionaries/spouse_dict_25Jun2020.txt', sep = ',', ncolumns = length(spouse))
```
