---
title: "Family Communication-Enriched Notes"
output: html_document
---

## Strategy

1. All ICU Admissions
2. Remove pts with missing CHARTEVENTS data
3. Remove pts younger than 18
4. Keep only pts who were mechanically ventilated during their admission
5. Remove pts without any notes associated with their admission or notes logged in error
6. Keep only pts with physicians' notes associated with their admission (2008-2012)

### Utility Functions

```{r, echo = F}
## Rename columns, avoiding some
rename <- function(dat, avoid_names, prefix){
        for (name in colnames(dat)[!(colnames(dat) %in% avoid_names)]){
            colnames(dat)[which(colnames(dat) == name)] <- paste(prefix, name, sep = '')
        }
    return(dat)
}

```

#### clean_text

```{r}
clean_text <- function(tokens, printout){
    #Create a fake patient note phrase as a representative sample
    ex_token <- "Example note:\nThe patient is a 81yo m who was found down in [** location **] on [** date **] by daughter, [** name **].\n Pt was in usual state of health until four days ago, when began to complain to family of heartburn for which the pt was taking tums in addition to his prescribed PPI, without resolution."
  if (printout){
    print(substr(ex_token, 1, 100))
  }
  
  #Remove carriage returns, convert to lower
  tokens <- tolower(gsub('\r', ' ', tokens))
  tokens <- tolower(gsub('\n', ' ', tokens))
  ex_token <- tolower(gsub('\n', ' ', ex_token))
  if (printout){
    cat("Result after removing carriage returns:\n")
    print(substr(ex_token, 1, 100))
  }
  
  #https://stackoverflow.com/questions/13529360/replace-text-within-parenthesis-in-r
  #Remove obfuscations between '[' and ']'
  tokens <- gsub(" *\\[.*?\\] *", ' ', tokens)
  ex_token <- gsub(" *\\[.*?\\] *", ' ', ex_token)
  if (printout){
    cat("Result after leaving [obfuscation]:\n")
    print(substr(ex_token, 1, 100))
  }
  
  #Keep only words & numeric
  tokens <- gsub("[^[:alnum:][:space:]]", '', tokens)
  ex_token <- gsub("[^[:alnum:][:space:]]", '', ex_token)
  if (printout){
    cat("Result after removing all but alphanumeric and spaces:\n")
    print(substr(ex_token, 1, 100))
  }
  
  #Keep only a single white space
  #https://stackoverflow.com/questions/25707647/merge-multiple-spaces-to-single-space-remove-trailing-leading-spaces
  tokens <- gsub("(?<=[\\s])\\s*|^\\s+|\\s+$", '', tokens, perl=TRUE)
  ex_token <- gsub("(?<=[\\s])\\s*|^\\s+|\\s+$", '', ex_token, perl=TRUE)
  if (printout){
    cat("Result after keeping only single spaces:\n")
    print(substr(ex_token, 1, 100))
  }
  
  ## Drop all whitespace
  tokens <- gsub(' ', '', tokens)
  
  return(tokens)
}
```

# Demographic and Cohort Data

## Load Admission Data & Initial Cleaning

Load `ADMISSIONS`, `PATIENTS`, and `ICUSTAYS` for cohort data. Load `icustay_detail` for more derived data. __Note:__ `icustay_detail` data are derived from `ICUSTAYS`

```{r}
## Load ADMISSIONS Table to join on HADM_ID
adm <- read.csv("~/MIMIC-III/ADMISSIONS.csv", 
                header = T, stringsAsFactors = F)

## Load PATIENTS Table to join on SUBJECT_ID
pat <- read.csv("~/MIMIC-III/PATIENTS.csv", 
                header = T, stringsAsFactors = F)

## Load ICUSTAYS Table to join on SUBJECT_ID, HADM_ID
stays <- read.csv("~/MIMIC-III/ICUSTAYS.csv", 
                  header = T, stringsAsFactors = F)

## Load icustay_detail to join with ICUSTAYS
## Note: icustay_detail is a derived table
stays_detail <- read.csv("~/MIMIC-III/icustay_detail.csv", 
                         header = T, stringsAsFactors = F)
## Convert colnames to uppercase
colnames(stays_detail) <- toupper(colnames(stays_detail))
```

## Initial Cohort (Derived from ICUSTAYS)

As all patients must be acutely ill, we will begin at the ICU level.

```{r}
cat("There are", length(unique(stays$SUBJECT_ID)), "unique patients in this cohort.\n")
cat("There are", length(unique(stays$HADM_ID)), "unique hospital admissions associated with this cohort.\n")
cat("There are", length(unique(stays$ICUSTAY_ID)), "unique ICUSTAYs in this cohort.\n")

## If they are not in the detailed stays table, it's because they are missing data
cat(length(unique(stays$SUBJECT_ID)) - length(unique(stays_detail$SUBJECT_ID)), "Patients were removed for having no CHARTEVENTS data.\n")
cat(length(unique(stays$HADM_ID)) - length(unique(stays_detail$HADM_ID)), "HOSPITAL ADMISSIONS are missing CHARTEVENTS data.\n")
cat(length(unique(stays$ICUSTAY_ID)) - length(unique(stays_detail$ICUSTAY_ID)), "ICUSTAYs are missing CHARTEVENTS data.\n")
```

### Merging Admission, Patient, and ICU Data

1. Remove `ROW_ID`, which is an index peculiar to each table.
2. Join `ICUSTAYS` to derived `ICUSTAYS_DETAIL` table on tabular intersection. `ICUSTAYS` contains information that the derived `ICUSTAYS_DETAIL` does not.
3. Join `ADMISSIONS` to `PATIENTS` on `SUBJECT_ID` for `GENDER`, `DOB`, `DOD` info.
4. Join `ICUSTAYS` to `PATIENTS` and `ADMISSIONS` on tabular intersection.

```{r}
## (1)
adm$ROW_ID <- NULL
pat$ROW_ID <- NULL
stays$ROW_ID <- NULL

## (2)
stays <- merge(stays, stays_detail, by = intersect(colnames(stays), colnames(stays_detail)))

## Clean stays_detail
rm(stays_detail)

## (3)
tmp <- merge(adm, pat, by = "SUBJECT_ID")
dim(tmp)

## Clean
rm(adm, pat)

## (4)
tmp <- merge(tmp, stays, by = intersect(colnames(tmp), colnames(stays)))
dim(tmp)

## Clean
rm(stays)

## Look again for pt, hadm, icustay numbers
cat("There are", length(unique(tmp$SUBJECT_ID)), "unique patients in this cohort.\n")
cat("There are", length(unique(tmp$HADM_ID)), "unique hospital admissions in this cohort.\n")
cat("There are", length(unique(tmp$ICUSTAY_ID)), "unique ICUSTAYs in this cohort.\n")

dat <- tmp
rm(tmp)
```

### Correct Nonagenarian Ages and Remove patients younger than 18 years of age

```{r}
## Correct nonagenarian ages by imputing median age (91.4)
dat[(dat$ADMISSION_AGE >= 90), ]$ADMISSION_AGE <- 91.4

## Remove patients youner than 18
tmp <- dat[(dat$ADMISSION_AGE >= 18), ]

cat(length(unique(dat$SUBJECT_ID)) - length(unique(tmp$SUBJECT_ID)), "patients below the age of 18 removed.\n")
cat(length(unique(dat$HADM_ID)) - length(unique(tmp$HADM_ID)), "hospital admissions for patients below the age of 18 removed.\n")
cat(length(unique(dat$ICUSTAY_ID)) - length(unique(tmp$ICUSTAY_ID)), "ICU stays for patients below the age of 18 removed.\n")

## dat gets tmp, remove tmp
dat <- tmp
rm(tmp)

cat("There are", length(unique(dat$SUBJECT_ID)), "unique patients in this cohort.\n")
cat("There are", length(unique(dat$HADM_ID)), "unique hospital admissions in this cohort.\n")
cat("There are", length(unique(dat$ICUSTAY_ID)), "unique ICUSTAYs in this cohort.\n")
```

## Check Ventilator Use

1. Load `ventdurations` (derived table)
2. Merge data, thus removing any patients which have no ventilation events

```{r}
## (1)
## ventdurations
vent <- read.csv("~/MIMIC-III/ventdurations.csv", header = T, stringsAsFactors = F)

## Convert column names to uppercase
colnames(vent) <- toupper(colnames(vent))

## (2)
## Remove patients with no ventilation data
tmp <- merge(dat, vent, by = "ICUSTAY_ID")

## Print numbers
cat(length(unique(dat$SUBJECT_ID)) - length(unique(tmp$SUBJECT_ID)), "patients removed for never being on a ventilator.\n")
cat(length(unique(dat$HADM_ID)) - length(unique(tmp$HADM_ID)), "hospital admissions removed for not being associated with ventilator use.\n")
cat(length(unique(dat$ICUSTAY_ID)) - length(unique(tmp$ICUSTAY_ID)), "ICU stays removed for not being associated with ventilator use.\n")

## dat gets tmp, remove tmp
dat <- tmp
rm(tmp, vent)

## Print
cat("There are", length(unique(dat$SUBJECT_ID)), "unique patients in this cohort.\n")
cat("There are", length(unique(dat$HADM_ID)), "unique hospital admissions in this cohort.\n")
cat("There are", length(unique(dat$ICUSTAY_ID)), "unique ICUSTAYs in this cohort.\n")
```

## Keep only MICU

```{r}
dat <- dat[dat$FIRST_CAREUNIT == "MICU", ]

cat("There are", length(unique(dat$SUBJECT_ID)), "unique patients in this cohort.\n")
cat("There are", length(unique(dat$HADM_ID)), "unique hospital admissions in this cohort.\n")
cat("There are", length(unique(dat$ICUSTAY_ID)), "unique ICUSTAYs in this cohort.\n")
```

## Notes

```{r}
## Load NOTEEVENTS table
notes <- read.csv("~/MIMIC-III/NOTEEVENTS.csv", header = T, stringsAsFactors = F)
```

### Join and Remove Notes

```{r}
tmp <- merge(dat, notes, by = c("SUBJECT_ID", "HADM_ID"))

## Clean environment
rm(notes)

## Remove notes marked ISERROR
tmp <- tmp[is.na(tmp$ISERROR), ]

## 
cat(length(unique(dat$SUBJECT_ID)) - length(unique(tmp$SUBJECT_ID)), "patients removed as not having notes or having marked errors.\n")
cat(length(unique(dat$HADM_ID)) - length(unique(tmp$HADM_ID)), "hospital admissions removed as not having notes or having marked errors.\n")
cat(length(unique(dat$ICUSTAY_ID)) - length(unique(tmp$ICUSTAY_ID)), "ICU stays removed as not having notes or having marked errors.\n")

## dat gets tmp, remove tmp
dat <- tmp
rm(tmp)

## Print
cat("There are", length(unique(dat$SUBJECT_ID)), "unique patients in this cohort.\n")
cat("There are", length(unique(dat$HADM_ID)), "unique hospital admissions in this cohort.\n")
cat("There are", length(unique(dat$ICUSTAY_ID)), "unique ICUSTAYs in this cohort.\n")
```

### Check and Subset Physician Notes (2008-2012)

Include social work/case management

```{r}
table(dat$CATEGORY)

## Subset pts with physicians' or Nursing notes
tmp <- dat[(dat$CATEGORY == "Physician " | dat$CATEGORY == "Nursing" | dat$CATEGORY == "Case Management " | dat$CATEGORY == "Social Work"), ]

## 
cat(length(unique(dat$SUBJECT_ID)) - length(unique(tmp$SUBJECT_ID)), "patients removed as not having Physicians' notes.\n")
cat(length(unique(dat$HADM_ID)) - length(unique(tmp$HADM_ID)), "hospital admissions removed as not having Physicians' notes.\n")
cat(length(unique(dat$ICUSTAY_ID)) - length(unique(tmp$ICUSTAY_ID)), "ICU stays removed as not having Physicians' notes.\n")

## dat gets tmp, remove tmp
dat <- tmp
rm(tmp)

## Print
cat("There are", length(unique(dat$SUBJECT_ID)), "unique patients in this cohort.\n")
cat("There are", length(unique(dat$HADM_ID)), "unique hospital admissions in this cohort.\n")
cat("There are", length(unique(dat$ICUSTAY_ID)), "unique ICUSTAYs in this cohort.\n")
```

## Cleaning

```{r}
## keep other variables in tmp for now
## Clean everything but alphanumeric from text
dat$STRIPPED_TEXT <- clean_text(dat$TEXT, TRUE)
```

## Remove Exact Duplicates

```{r}
tmp <- dat[!(duplicated(dat$STRIPPED_TEXT)),]

## 
cat(length(unique(dat$SUBJECT_ID)) - length(unique(tmp$SUBJECT_ID)), "patients removed as duplicate-note having.\n")
cat(length(unique(dat$HADM_ID)) - length(unique(tmp$HADM_ID)), "hospital admissions removed as duplicate-note having.\n")
cat(length(unique(dat$ICUSTAY_ID)) - length(unique(tmp$ICUSTAY_ID)), "ICU stays removed as duplicate-note having.\n")

## dat gets tmp, remove tmp
dat <- tmp
rm(tmp)

## Remove
dat$STRIPPED_TEXT <- NULL

## Print
cat("There are", length(unique(dat$SUBJECT_ID)), "unique patients in this cohort.\n")
cat("There are", length(unique(dat$HADM_ID)), "unique hospital admissions in this cohort.\n")
cat("There are", length(unique(dat$ICUSTAY_ID)), "unique ICUSTAYs in this cohort.\n")
cat("There are", length(unique(dat$TEXT)), "unique TEXTS in this cohort.\n")
```

## CHARTTIME_HRS_SINCE_ADMIT

```{r}
## CHARTTIME - ADMITTIME
dat$CHART_HRS_SINCE_ADMIT <- (as.numeric(strptime(dat$CHARTTIME, "%Y-%m-%d %H:%M:%S", tz = "EST")) - 
                                as.numeric(strptime(dat$ADMITTIME, "%Y-%m-%d %H:%M:%S", tz = "EST")))/60/60
```

### Additional Tables

#### Renal Replacement Therapy on Day One

```{r}
## Hold dat in tmp for now
tmp <- dat

rrtd1 <- read.csv("~/MIMIC-III/rrtfirstday.csv", header = T, stringsAsFactors = F)
colnames(rrtd1) <- toupper(colnames(rrtd1))
rrtd1 <- aggregate(RRT ~ ICUSTAY_ID, data = rrtd1, FUN = max)
colnames(rrtd1) <- c("ICUSTAY_ID", "RRT_D1")

## Merge
tmp <- merge(x = tmp, y = rrtd1, by = "ICUSTAY_ID", all.x = TRUE)

rm(rrtd1)
```

#### Sequential Organ Failure Assessment (SOFA)

```{r}
sofa <- read.csv("~/MIMIC-III/sofa.csv", header = T, stringsAsFactors = F)
colnames(sofa) <- toupper(colnames(sofa))

## Subset
sofa <- sofa[, c("SUBJECT_ID", "HADM_ID", "ICUSTAY_ID", "SOFA")]
sofa <- rename(sofa, c("SUBJECT_ID", "HADM_ID", "ICUSTAY_ID"), "ADMISSION_")

## Merge
tmp <- merge(x = tmp, y = sofa, by = c("SUBJECT_ID", "HADM_ID", "ICUSTAY_ID"), all.x = TRUE)
rm(sofa)
```

#### Elixhauser Comorbidities

```{r}
elixhauser <- read.csv("~/MIMIC-III/elixhauser.csv", header = T, stringsAsFactors = F)
colnames(elixhauser) <- toupper(colnames(elixhauser))
elixhauser <- rename(elixhauser, "HADM_ID", "ADMISSION_")

## Merge
tmp <- merge(x = tmp, y = elixhauser, by = "HADM_ID", all.x = TRUE)

## Clean
rm(elixhauser)
```

#### Vasopressors

```{r}
vaso <- read.csv("~/MIMIC-III/vasopressordurations.csv", header = T, stringsAsFactors = F)
colnames(vaso) <- toupper(colnames(vaso))
vaso <- rename(vaso, "ICUSTAY_ID", "VASOPRESSOR_")
## Drop vasopressor count
vaso$VASONUM <- NULL

## Sum vasopressor time
vaso <- aggregate(VASOPRESSOR_DURATION_HOURS ~ ICUSTAY_ID, data = vaso, FUN = sum)

## Merge back to data
tmp <- merge(x = tmp, y = vaso, by = c("ICUSTAY_ID"), all.x = TRUE)

## Clean
rm(vaso)
```

### Write

```{r}
# write.csv(tmp, file = "~/Desktop/CG_Data/caregivers_micu_noteevents02Jun20.csv", row.names = F)
```

