---
title: "Caregivers Regression Analysis"
output: html_document
---

# Overall Questions:

### Helper Functinos

```{r}
library(sjPlot)
library(sjstats)

model_info <- function(fit){
  model_sum <- summary(fit)
  odds_ratio <- cbind(OR = exp(fit$coef), exp(confint(fit)))
  my_list <- list(model_sum, odds_ratio)
  names(my_list) <- c("Model Summary", "OR Summary")
  return(my_list)
}

## to_hours
## convert MIMIC-III date time format to hours
to_hours <- function(some_time){
  return(as.numeric(strptime(some_time, "%Y-%m-%d %H:%M:%S", tz = "EST"))/(60*60))
}

## Factor data which have a few levels
factor_data <- function(dat){
    ## Any row with levels less than the log of the number of rows
    cols_to_factor <- sapply(dat, function(col) length(unique(col)) < log10(length(col) * 1000) )
    dat[cols_to_factor] <- lapply(dat[cols_to_factor], factor)
    return(dat)
}


```


## Load data, Rename some Columns

```{r}
dat <- read.csv('~/caregivers/data/processed/investigational_regex_results03Jun2020.csv', header = T, stringsAsFactors = F)

## Rename to avoid the term "ADMISSION" which will be used elsewhere
colnames(dat)[which(colnames(dat) == "ADMISSION_AGE")] <- "AGE"
colnames(dat)[which(colnames(dat) == "ADMISSION_SOFA")] <- "SOFA_SCORE"
colnames(dat)[which(colnames(dat) == "ADMISSION_TYPE")] <- "ADMIT_TYPE"
colnames(dat)[which(colnames(dat) == "ADMISSION_LOCATION")] <- "ADMIT_LOCATION"

## Rename "DURATION HOURS" to refer to mechvent
colnames(dat)[which(colnames(dat) == "DURATION_HOURS")] <- "VENT_DURATION_HOURS"
## Drop old data for sanity's sake
dat$VENT_DURATION_HOURS <- NULL

## Vasopressors-- correct NA to 0
dat$VASOPRESSOR_DURATION_HOURS <- ifelse(is.na(dat$VASOPRESSOR_DURATION_HOURS), 0, dat$VASOPRESSOR_DURATION_HOURS)
```

## Data Cleaning

### Create First 48 Variables

```{r}
SPOUSE_FIRST_48 <- vector()
CHILD_FIRST_48 <- vector()

for (i in 1:nrow(dat)){
  if ((dat$SPOUSE[i] == 1) & (dat$CHART_HRS_SINCE_ADMIT[i] <= 48)){
    SPOUSE_FIRST_48 <- c(SPOUSE_FIRST_48, 1)
  } else {
    SPOUSE_FIRST_48 <- c(SPOUSE_FIRST_48, 0)
  }
  
  if ((dat$CHILD[i] == 1) & (dat$CHART_HRS_SINCE_ADMIT[i] <= 48)){
    CHILD_FIRST_48 <- c(CHILD_FIRST_48, 1)
  } else {
    CHILD_FIRST_48 <- c(CHILD_FIRST_48, 0)
  }
}

dat <- cbind(dat, SPOUSE_FIRST_48, CHILD_FIRST_48)

## Clean environment
rm(SPOUSE_FIRST_48, CHILD_FIRST_48, i)
```

## Create mortality info

```{r}
## Generate date data in DAYS
dat$DAYS_UNTIL_DEATH_DISCH <- (to_hours(dat$DOD) - to_hours(dat$DISCHTIME))/24

## Correct negatives (in-hospital mortality)
dat$DAYS_UNTIL_DEATH_DISCH <- ifelse(dat$DAYS_UNTIL_DEATH_DISCH < 0, 0, dat$DAYS_UNTIL_DEATH_DISCH)

## Correct for those who didn't pass away-- correct to maximum survival time
dat$DAYS_UNTIL_DEATH_DISCH <- ifelse(is.na(dat$DAYS_UNTIL_DEATH_DISCH), max(dat$DAYS_UNTIL_DEATH_DISCH, na.rm = T), dat$DAYS_UNTIL_DEATH_DISCH)

dat$MORT_3MO <- ifelse(dat$DAYS_UNTIL_DEATH_DISCH <= 90, 1, 0)
dat$MORT_3MO <- ifelse(is.na(dat$MORT_3MO), 0, dat$MORT_3MO)

dat$MORT_1YR <- ifelse(dat$DAYS_UNTIL_DEATH_DISCH <= 365, 1, 0)
dat$MORT_1YR <- ifelse(is.na(dat$MORT_1YR), 0, dat$MORT_1YR)
```

## Level Changes for Certain Data

```{r}
## Reduce the number of ethnicities by scoping outward
dat[(grepl("WHITE", dat$ETHNICITY)),]$ETHNICITY <- "WHITE"
dat[(grepl("PORTUGUESE", dat$ETHNICITY)),]$ETHNICITY <- "WHITE"
dat[(grepl("BLACK", dat$ETHNICITY)),]$ETHNICITY <- "BLACK"
dat[(grepl("HISPANIC", dat$ETHNICITY)),]$ETHNICITY <- "HISPANIC"
dat[(grepl("ASIAN|HISPANIC|MIDDLE|NATIVE|MULTI|DECLINED|UNABLE|OTHER|NOT", dat$ETHNICITY)),]$ETHNICITY <- "UNKNOWN/OTHER"

## Marriage
dat$MARITAL_STATUS <- ifelse(dat$MARITAL_STATUS == "WIDOWED", "SEPARATED", dat$MARITAL_STATUS)
dat$MARITAL_STATUS <- ifelse(dat$MARITAL_STATUS == "DIVORCED", "SEPARATED", dat$MARITAL_STATUS)
dat$MARITAL_STATUS <- ifelse(dat$MARITAL_STATUS == "", "SINGLE/UNK", dat$MARITAL_STATUS)
dat$MARITAL_STATUS <- ifelse(dat$MARITAL_STATUS == "SINGLE", "SINGLE/UNK", dat$MARITAL_STATUS)

## Change levels for tabulation
## Admission Location
dat[(grepl("REFERRAL", dat$ADMIT_LOCATION)),]$ADMIT_LOCATION <- "PHYS/CLINIC REFERRAL"
dat[(grepl("TRANSFER", dat$ADMIT_LOCATION)),]$ADMIT_LOCATION <- "TRANSFER FROM HOSP/SNF"

## Discharge Location
## Rehab = SNF
dat$DISCHARGE_LOCATION <- ifelse(dat$DISCHARGE_LOCATION == "REHAB/DISTINCT PART HOSP", "SNF", dat$DISCHARGE_LOCATION)
## Aggregate SHORT TERM HOSPITAL and OTHER
dat$DISCHARGE_LOCATION <- ifelse(dat$DISCHARGE_LOCATION == "SHORT TERM HOSPITAL", "SHORT TERM HOSPITAL/OTHER", dat$DISCHARGE_LOCATION)
dat$DISCHARGE_LOCATION <- ifelse(dat$DISCHARGE_LOCATION == "OTHER FACILITY", "SHORT TERM HOSPITAL/OTHER", dat$DISCHARGE_LOCATION)

## Hospice
dat[(grepl("HOSPICE", dat$DISCHARGE_LOCATION)),]$DISCHARGE_LOCATION <- "HOSPICE"
## DISC ~= HOME HEALTH
dat[(grepl("DISC", dat$DISCHARGE_LOCATION)),]$DISCHARGE_LOCATION <- "HOME HEALTH CARE"
## Home = Home/Home Health
dat[(grepl("HOME", dat$DISCHARGE_LOCATION)),]$DISCHARGE_LOCATION <- "HOME/HOME HEALTH CARE"
## Left = Home
dat[(grepl("LEFT", dat$DISCHARGE_LOCATION)),]$DISCHARGE_LOCATION <- "HOME/HOME HEALTH CARE"

## Bucket languages to ENGL/Other
dat$LANGUAGE <- ifelse(dat$LANGUAGE == "ENGL", dat$LANGUAGE, "OTHER")

## Elixhauser
dat$ELIXHAUSER_SCORE <- apply(dat[,grepl("ADMISSION_", colnames(dat))], 1, sum)
```

### Drop Unnecessary Data

```{r}
dat <- dat[,!(colnames(dat) %in% c("ADMITTIME",
               "DISCHTIME",
               "DOD",
               "DEATHTIME",
               "EDREGTIME",
               "EDOUTTIME",
               "DIAGNOSIS",
               "HAS_CHARTEVENTS_DATA",
               "DOB",
               "DOD_HOSP",
               "DOD_SSN",
               "EXPIRE_FLAG",
               "INTIME",
               "OUTTIME",
               "DBSOURCE",
               "ROW_ID",
               "LAST_CAREUNIT",
               "FIRST_WARDID",
               "LAST_WARDID",
               "LOS",
               "INSURANCE",
               "RELIGION",
               "VENTNUM",
               "STARTTIME",
               "ENDTIME",
               "CHARTDATE",
               "CHARTTIME",
               "STORETIME",
               "TEXT",
               "DESCRIPTION",
               ## Maybe keep category..
               "CATEGORY",
               "CGID",
               "CHILD",
               "SPOUSE",
               "ISERROR"))]
```

## Ventilation Time

Create `PERC_VENT_DUR` = `sum(VENT_DURATION)`/`ICU_LOS`

1. Aggregate `VENT_DURATION_HOURS` by HOSPITAL ADMISSION (same `n` as ICU ADMISSION)


```{r}
## Reload durations data
vent <- read.csv("~/MIMIC-III/ventdurations.csv", header = T, stringsAsFactors = F)
colnames(vent) <- toupper(colnames(vent))

## Vent has multiple durations for ICUSTAY's, aggregate them
vent <- aggregate(DURATION_HOURS ~ ICUSTAY_ID, data = vent, FUN = sum)

## Don't remerge yet
## dat <- merge(..., vent, by = "ICUSTAY_ID", all.x = T)

los <- dat[,c("ICUSTAY_ID", "LOS_ICU", "LOS_HOSPITAL")]

## Remove duplicates
los <- los[!duplicated(los$ICUSTAY_ID),]

## Merge
vent <- merge(los, vent, by = "ICUSTAY_ID")

## Clean
rm(los)

## Create Duration Days
vent$DURATION_DAYS <- vent$DURATION_HOURS/24
## PERC_ICU_VENT
vent$PERC_ICU_VENT <- vent$DURATION_DAYS/vent$LOS_ICU*100

head(vent)

## Drop unnecessary variables
vent$LOS_HOSPITAL <- NULL
vent$LOS_ICU <- NULL
```

### Aggregate to Admission Level

```{r}
tmp <- aggregate(cbind(CHILD_FIRST_48,
                       SPOUSE_FIRST_48,
                       MORT_1YR,
                       MORT_3MO,
                       HOSPITAL_EXPIRE_FLAG,
                       AGE,
                       VASOPRESSOR_DURATION_HOURS,
                       LOS_HOSPITAL,
                       LOS_ICU,
                       HOSPSTAY_SEQ,
                       RRT_D1,
                       ELIXHAUSER_SCORE,
                       SOFA_SCORE,
                       ADMISSION_CONGESTIVE_HEART_FAILURE, 
                       ADMISSION_CARDIAC_ARRHYTHMIAS,
                       ADMISSION_VALVULAR_DISEASE,
                       ADMISSION_PULMONARY_CIRCULATION, 
                       ADMISSION_PERIPHERAL_VASCULAR, 
                       ADMISSION_HYPERTENSION, 
                       ADMISSION_PARALYSIS, 
                       ADMISSION_OTHER_NEUROLOGICAL, 
                       ADMISSION_CHRONIC_PULMONARY, 
                       ADMISSION_DIABETES_UNCOMPLICATED, 
                       ADMISSION_DIABETES_COMPLICATED, 
                       ADMISSION_HYPOTHYROIDISM, 
                       ADMISSION_RENAL_FAILURE, 
                       ADMISSION_LIVER_DISEASE, 
                       ADMISSION_PEPTIC_ULCER, 
                       ADMISSION_AIDS, 
                       ADMISSION_LYMPHOMA, 
                       ADMISSION_METASTATIC_CANCER, 
                       ADMISSION_SOLID_TUMOR, 
                       ADMISSION_RHEUMATOID_ARTHRITIS, 
                       ADMISSION_COAGULOPATHY, 
                       ADMISSION_OBESITY, 
                       ADMISSION_WEIGHT_LOSS, 
                       ADMISSION_FLUID_ELECTROLYTE, 
                       ADMISSION_BLOOD_LOSS_ANEMIA, 
                       ADMISSION_DEFICIENCY_ANEMIAS, 
                       ADMISSION_ALCOHOL_ABUSE, 
                       ADMISSION_DRUG_ABUSE, 
                       ADMISSION_PSYCHOSES, 
                       ADMISSION_DEPRESSION) ~
                       GENDER +
                   ETHNICITY +
                   LANGUAGE +
                   MARITAL_STATUS +
                   FIRST_CAREUNIT +
                   FIRST_ICU_STAY +
                   ADMIT_LOCATION +
                   DISCHARGE_LOCATION +
                   SUBJECT_ID + 
                   HADM_ID +
                   ICUSTAY_ID, data = dat, FUN = max)
```

### Merge to ventilation data

```{r}
temp <- merge(tmp, vent, by = "ICUSTAY_ID")
```


```{r}
## Write
write.csv(temp, file = "~/caregivers/data/processed/regression_data11Jun2020.csv", row.names = F)
```
